---
output: github_document
---

```{r setup, echo = FALSE}
knitr::opts_chunk$set(
  warning = FALSE,
  echo = TRUE,
  collapse = TRUE,
  message = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)
```

# epimetrics

<!-- badges: start -->
[![R-CMD-check](https://github.com/NicChr/epimetrics/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/NicChr/epimetrics/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/NicChr/epimetrics/branch/main/graph/badge.svg)](https://app.codecov.io/gh/NicChr/epimetrics?branch=main)
<!-- badges: end -->

## A package for easy calculation of diagnostic test metrics with confidence intervals

## Installation

```{r gh-installation, message = FALSE, eval = FALSE}
remotes::install_github("NicChr/epimetrics")
```

```{r package_load}
library(epimetrics)
```

You can easily calculate metrics from data, which is typically
compared to a gold standard

```{r confusion}
# Using a 2x2 confusion matrix example from wikipedia 
# https://en.wikipedia.org/wiki/Sensitivity_and_specificity
TP <- 20
FP <- 180
FN <- 10
TN <- 1820
CM <- matrix(c(TP, FP, FN, TN), byrow = TRUE, ncol = 2)
rownames(CM) <- c("+Test", "-Test")
colnames(CM) <- c("+Disease", "-Disease")
CM
set.seed(42)
epimetrics(CM)
```

0r from known or theoretical estimates of sensitivity, specificity 
and prevalence

```{r predictive_value}
sens <- 0.75 # Sensitivity
spec <- 0.95 # Specificity
prev <- seq(0, 1, 0.01) # Prevalence
ppv <- ppv(sens, spec, prev)
npv <- npv(sens, spec, prev)
plot(prev, ppv, col = "blue")
points(prev, npv, col = "orange")
legend(0.6, 0.4, legend = c("PPV", "NPV"),
       col = c("blue", "orange"),
       pch = c(1, 1))
```

You can also use standalone metrics

```{r standalone}
sensitivity(x = CM)
specificity(x = CM)
ppv(x = CM, conf.level = 0.95)
```

Easily simulate confusion matrix data

```{r simulation}
simulated_data <- simulate_confusion_data(TP, FP, FN, TN, R = 10)
simulated_data
ppv(x = simulated_data, conf.int = FALSE)
ppv(x = simulated_data)
```

If your data are in non-aggregate form, you can use `confusion_matrix`

```{r confusion2}
outcomes <- c(rep(1, TP + FN), # With disease
              rep(0, FP + TN)) # Without disease
predictions <- c(rep(1, TP), rep(0, FN), # Disease group predictions
                 rep(1, FP), rep(0, TN)) # Non-disease group predictions
(cm <- confusion_matrix(outcomes, predictions))
set.seed(42)
epimetrics(cm)
```



